---
title: "P8105_hw3_jl5172"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load package and data cleaning
```{r}
library(tidyverse)
library(p8105.datasets)
data("brfss_smart2010")
brfss<-janitor::clean_names(brfss_smart2010) #clean names
brfss<-rename(brfss,state=locationabbr,location=locationdesc) # use appropriate variable names
brfss<-filter(brfss,topic == "Overall Health") # fiocusing on Overall Health
brfss<-filter(brfss,response=="Excellent"|response=="Very good"|response=="Good"|response=="Fair"|response=="Poor") #Keep responses required
mutate(brfss,response=as.factor(brfss$response)) 
brfss$response<-fct_relevel(brfss$response,"Excellent","Very good","Good","Fair","Poor") # organize response as facctors in required ordering

```

In 2002, which states were observed at 7 locations?
```{r}
brfss %>% 
  filter(year=="2002") %>%   #filter by year 2002
  distinct(state,location) %>%  # distinct by state and location
  count(state) %>%  # count the number of observations for each states
  filter(n==7)  # looking for states which have been observed for 7 locations
```
We can see that CT,FL and NC were observed at 7 location in 2002.

Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.

```{r}
brfss %>% 
  group_by(state,year) %>%  # group by state and year
  distinct(location) %>%   #distinct by locations
  count(state) %>%  #count how many locations in each state were observed in each year
  ggplot(aes(x = year, y = n)) +  #use ggplot to plot spaghetti
    geom_line(aes(color = state))+  # color each line by state 
  labs(
    title = "Number Of Locations in Each States 2002-2010",
    x = "Year",
    y = "Number of Locations",
    caption = "Data from the p8105.datasets package"
  ) +
  viridis::scale_color_viridis(
    name = "", 
    discrete = TRUE)+
  theme(legend.position = "right")
```



Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

```{r}
brfss %>% 
  filter(year=="2002"|year=="2006"|year=="2010",state=="NY",response=="Excellent") %>% 
  group_by(year) %>% 
  summarize(
          mean_proportion_excellent_NY =round(mean(data_value)/100,3),
          sd_excellent_response=round(sd(data_value),3)) %>% 
  View()
```
We can see from the table that...



For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.
```{r}
brfss %>% 
  spread(key= response, value = data_value) %>% 
  janitor::clean_names() %>% 
  group_by(state,year) %>% 
  summarize(excellent_mean=mean(excellent,na.rm = T),
            verygood_mean=mean(very_good,na.rm = T),
            good_mean=mean(good,na.rm=T),
            fair_mean=mean(fair,na.rm=T),
            poor_mean=mean(poor,na.rm=T)
            ) %>% 
 gather(key=response_variable,value=mean_value,excellent_mean:poor_mean) %>% 
  ggplot(aes(x=year,y=mean_value,color=state))+
  geom_line()+
  facet_grid(~response_variable)+
  labs(
    title = "Average Proportion In Each Response Category",
    x = "Mean Proportion",
    y = "Year",
    caption = "Data from the p8105.datasets package"
  )
```

